Application of Generalized Additive Models for long-term and seasonal trend analysis of water quality: South San Francisco Bay case study

### 2020-06-09 manu_draft.Rmd
trends example plot added

data(locs)
data(modprf)
data(seastrnd)
data(mods_chl18)
data(mods_chl24)
data(mods_chl27)
data(mods_chl32)

@@

Table \@ref(tab:modprftab), Table \@ref(tab:seastrndtab), Figure \@ref(fig:prddat)

@@

>>>>

data(locs)
data(modprf)
data(seastrnd)
data(mods_chl18)
data(mods_chl24)
data(mods_chl27)
data(mods_chl32)
data(mods_chl34)

@@

Table \@ref(tab:modprftab), Table \@ref(tab:seastrndtab), Figure \@ref(fig:prddat), Figure \@ref(fig:trnddat)

@@

```{r, results = 'hide', message = F}
toplo <- mods_chl34 %>% 
  filter(model %in% 'gam6') %>% 
  pull(modi)

ylim <- c(-1, 19)
ylab <- expression(paste("Chlorophyll-a (", mu, "g ", L^-1, ")"))

p1 <- show_perchg(mods = toplo, baseyr = 1991, testyr = 2000, ylab = ylab) +
  labs(
    subtitle = '(a) 1990, 2000, Base: 3.6, Test: 5.7,\nChange: 57.7%, ns', 
    title = NULL
  )
p2 <- show_perchg(mods = toplo, baseyr = 2000, testyr = 2010, ylab = NULL) +
  labs(
    subtitle = '(b) 2000, 2010, Base: 5.7, Test: 10.1,\nChange: 77.5%, p < 0.001', 
    title = NULL
  )
p3 <- show_perchg(mods = toplo, baseyr = 2010, testyr = 2017, ylab = NULL) +
  labs(
    subtitle = '(c) 2010, 2017, Base: 10.1, Test: 6.1,\nChange: -39.8%, p < 0.01', 
    title = NULL
  )
p4 <- show_avgseason(mods = toplo, doystr = 41, doyend = 213, yrstr = 1991, yrend = 2000, ylab = ylab) +
  scale_y_continuous(limits = ylim) + 
  labs(
    subtitle = '(d) 1990-2000, Jan-Jul, slope 0.6, p < 0.01', 
    title = NULL
  )
p5 <- show_avgseason(mods = toplo, doystr = 41, doyend = 213, yrstr = 2000, yrend = 2010, ylab = NULL) +
  scale_y_continuous(limits = ylim) + 
  labs(
    subtitle = '(e) 2000-2010, Jan-Jul, slope 0.5, ns', 
    title = NULL
  )
p6 <- show_avgseason(mods = toplo, doystr = 41, doyend = 213, yrstr = 2010, yrend = 2016, ylab = NULL) +
  scale_y_continuous(limits = ylim) + 
  labs(
    subtitle = '(f) 2010-2017, Jan-Jul, slope -0.4, ns', 
    title = NULL
  )
p7 <- show_avgseason(mods = toplo, doystr = 213, doyend = 338, yrstr = 1991, yrend = 2000, ylab = ylab) +
  scale_y_continuous(limits = ylim) + 
  labs(
    subtitle = '(g) 1990-2000, Aug-Dec, slope 0.26, ns', 
    title = NULL
  )
p8 <- show_avgseason(mods = toplo, doystr = 213, doyend = 338, yrstr = 2000, yrend = 2010, ylab = NULL) +
  scale_y_continuous(limits = ylim) + 
  labs(
    subtitle = '(h) 2000-2010, Aug-Dec, slope 0.56, p < 0.05', 
    title = NULL
  )
p9 <- show_avgseason(mods = toplo, doystr = 213, doyend = 338, yrstr = 2010, yrend = 2017, ylab = NULL) +
  scale_y_continuous(limits = ylim) + 
  labs(
    subtitle = '(i) 2010-2017, Aug-Dec, slope -0.52, ns', 
    title = NULL
  )

p <- p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 + p9 + plot_layout(nrow = 3) & 
  theme_minimal(base_family = 'serif', base_size = 12) &
  theme(
    axis.title.x = element_blank(),
    legend.position = 'none'
    )

png('figs/trnddat.png', height = 9, width = 11, family = 'serif', units = 'in', res = 400)
print(p)
dev.off()
```
```{r trnddat, fig.cap = "Examples of trend tests using results of GAM predictions for station 34 (`gam6`).  Plots (a), (b), and (c) show estimates of percent change as the difference of means between base and test years for chlorophyll.  Plots (d) through (i) show seasonal averages and trend estimates over time.  Plot (d), (e), and (f) show trend estimates for January through July and (g), (h), and (i) show trend estimates for August through December.  The trend lines in (d) through (i) estimate the rate of change of chlorophyll per year, reported as the slope in the plot title. ns: not significant"}
knitr::include_graphics('figs/trnddat.png')
```
